Parallel Global Search Algorithm for Optimization of the Kinetic Parameters of Chemical Reactions

### 2021-05-25 2021/RSD_Ufa/RSD2021.tex
add three new references

A method for solving the optimization problems of such type should account for the following aspects. First, the objective function of the problem is a ``black box'' since it is defined not by a formula but in the form of some algorithm for computing its values. Second, computing even a single value of the objective function is time-consuming operation since it requires numerical modeling. The aspects listed above don't allow employing the multistart methods based on the ideas of multiple gradient descents for solving the inverse problems. On one hand, computing the gradients would be too expensive operation in this case. On the other hand, the multistart methods are hardly applicable for the essentially multiextremal problems, which the inverse problems of chemical kinetics belong to.

@@

The parallel global search algorithm described in this section is based on the idea of reducing the dimension with the use of the Peano curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}). By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
\]
where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
\]
with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
$ H=2 L \sqrt{N+3}$ and $y(x)$ is a Peano curve from $[0,1]$ onto $D$.

>>>>

A method for solving the optimization problems of such type should account for the following aspects. First, the objective function of the problem is a ``black box'' since it is defined not by a formula but in the form of some algorithm for computing its values. Second, computing even a single value of the objective function is time-consuming operation since it requires numerical modeling \cite{Modorskii2016}. The aspects listed above don't allow employing the multistart methods based on the ideas of multiple gradient descents for solving the inverse problems. On one hand, computing the gradients would be too expensive operation in this case. On the other hand, the multistart methods are hardly applicable for the essentially multiextremal problems, which the inverse problems of chemical kinetics belong to.

@@

Many state-of-the-art global optimization algorithms use the idea of reducing the dimensionality for solving multidimensional problems; see, for example, the diagonal partitions method \cite{Sergeyev2006} or the simplicial partitions method \cite{Zilinskas2008}. The parallel global search algorithm described in this section uses the Peano curves \cite{Sergeyev2013,Strongin2000}, which continuously and unambiguously map the unit interval $[0,1]$ onto the $N$-dimensional cube $D$ from (\ref{D}). By using this kind of mapping, it is possible to reduce the multidimensional problem (\ref{problem}) to a univariate problem
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
\]
where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
\]
with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation
$ H=2 L \sqrt{N+3}$ and $y(x)$ is a Peano curve from $[0,1]$ onto $D$.
